---
title: "Lab 5"
author: "Lizzy Schattle"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#attach packages
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(sf)
library(tmap)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(mapview)
library(paletteer)
```

## Monthly US Energy Consumption (Renewables)

```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

```

- Make description all lowercase
- only keep observations where description variable contains "consumption"
- remove any observations where description variable contains "total"

```{r}
renew_clean <- us_renew %>%
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>%
  filter(!str_detect(description, pattern = "total")) # ! means do the opposite of what you ask
```

### Convert 'yyyymm' column to date

```{r}
renew_date <- renew_clean %>%
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# make version where month and year are in separate columns:
renew_parsed <-  renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

### look back at it

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep,
                                          y = value,
                                          group = description,
                                          color = description)) +
  geom_line() +
  theme_bw()

renew_gg
```

Updating colors with paleteer palettes:

```{r}
renew_gg +
  scale_color_paletteer_d("nationalparkcolors::Badlands")
```


### coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```


look at time series data in a couple ways:

```{r}
renew_ts %>% 
  autoplot(value)

renew_ts %>% 
  gg_subseries(value)

# renew_ts %>% 
  # gg_season(value)


ggplot(data = renew_parsed,
       aes(x = month,
           y = value,
           group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right")
  
```


### Just look at hydroelectric energy consumption

```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% 
  autoplot(value)

hydro_ts %>% 
  gg_subseries(value)

#hydro_ts %>% 
  #gg_season(value)


ggplot(hydro_ts, aes(x = month,
                     y = value,
                     group = year)) +
  geom_line(aes(color = year))
```

### If i want the quarterly average consumption for hydroelectric power

```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu= ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```



### decompose that hydro_ts

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% 
  autoplot()
```


```{r}
hist(components(dcmp)$remainder)
```

now we look at ACF
```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```
### DANGER FORECASTING!!!!!
```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value),
    ETS(value)
  ) %>%
  fabletools::forecast(h = "4 years")

hydro_model %>% 
  autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

### make a world map woohoo!!
```{r}
world <- read_sf(here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

#mapview(world)
```








